--- 
title: "Reproducible Research: Peer Assessment 1" 
output: 
  html_document: 
    keep_md: true 
---


## Loading and preprocessing the data
```{r read_file}
DF <- read.csv("activity.csv", header=T, sep=',', na.strings="NA", check.names=F, stringsAsFactors=F, comment.char="", quote='\"')
DF2 <- DF
DF2$Date <- as.Date(as.character(DF2$date,format="%Y-%m-%d"))
```

## What is mean total number of steps taken per day?
```{r process_data}
DF2$steps[is.na(DF2$steps)] <- 0
cat(paste("Total number of actiivities taken ",nrow(DF2)))
print ("\n")
cat(paste("Mean steps taken per day ",round(mean(DF2$steps),1)))
print ("\n")
cat(paste("Median steps taken per day ",median(DF2$steps)))
print ("\n")
library(dplyr)
tot_steps <- aggregate(DF2$steps, by=list(date=DF2$date), FUN=sum)
names(tot_steps)[2]<-paste("total_steps")
tot_steps$date <- as.Date(as.character(tot_steps$date,format="%Y-%m-%d"))
with (tot_steps, hist(date, breaks = 61,  main="Total Number of Step Taken Each Day", ylab="Total Steps", xlab="Day", col="green"))
```

## What is the average daily activity pattern?
```{r average_daily_patten}
mean_steps <- aggregate(DF2$steps, by=list(date=DF2$date), FUN=mean)
mean_steps$date <- as.Date(as.character(mean_steps$date,format="%Y-%m-%d"))
with (mean_steps, plot(date, x, pch=20, type = "l"))
abline(h=median(mean_steps$x))
title(main="Mean Steps per Day", ylab="Average Number of Steps")
```

## Imputing missing values
There are
```{r} nrow(filter(DF, is.na(steps)))
```
  rows with missing values.
This data set has
```{r} nrows(DF)
```
 nows of data.

```{r plot_data}
DF3 <- DF
DF3$Date <- as.Date(as.character(DF3$date,format="%Y-%m-%d"))
DF3$steps[is.na(DF3$steps)] <- median(DF3$steps,na.rm = TRUE )
tot_steps3 <- aggregate(DF3$steps, by=list(date=DF3$date), FUN=sum)
names(tot_steps3)[2]<-paste("total_steps")
tot_steps3$date <- as.Date(as.character(tot_steps3$date,format="%Y-%m-%d"))
with (tot_steps3, hist(date, breaks = 61,  main="Total Number of Step Taken Each Day", ylab="Derived Total Steps", xlab="Day", col="red"))
```
Replacing the missing values does NOT impact the mean and median total number of steps taken per day.

```{r}
mean_steps3 <- aggregate(DF3$steps, by=list(date=DF3$date), FUN=mean)
mean_steps3$date <- as.Date(as.character(mean_steps3$date,format="%Y-%m-%d"))
par(mfrow = c(1, 1), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
with (mean_steps3, plot(date, x, xlab = "Date", ylab="Derived Average Number of Steps",main = "Derived Mean Steps per Day", type="l", col="red"))
abline(h=median(mean_steps3$x))
```

The both graphs have the same plots.

## Are there differences in activity patterns between weekdays and weekends?

```{r find_DOW_pattern}
DF3$DOW <- as.POSIXlt(DF3$date)$wday
DOW_steps <- aggregate(DF3$steps, by=list(DOW=DF3$DOW), FUN=sum)
DOW_steps$DOW[7] <- -1

barplot(DOW_steps$x, col="blue", ylim=c(0, 90000))
barplot(log10(DOW_steps$x), col="blue", ylim=c(0, 6))
axis(1, at=1:7, label=c("Sat", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri"))
mtext(side=1,text="Day of the Week",line=2.2)
mtext(side=2,text="Derived Steps per Day of the Week (log10)",line=2.2)
title(main="Total Steps per Day of the Week", col.main="red", font.main=4)
```

The log10(total steps) for weekends is ```{r} log10(DOW_steps$x[1] + DOW_steps$x[7]) ```
The log10(total steps) for weekdays is ```{r} log10(DOW_steps$x[2] + DOW_steps$x[3]+  DOW_steps$x[4] +  DOW_steps$x[5] +  DOW_steps$x[6])) ```
People are more active on weekends vs. weekdays.
 

